(n,p) Cross-Section Measurements for \ce{^{64}Cu} and \ce{^{47}Sc} via D-D Neutron Generator\\ OR \\(n,p) Cross-Section Measurements for \ce{^{64}Cu} and \ce{^{47}Sc} at 2.76 MeV using a D-D Neutron Generator for Medical Isotope Studies

### 2016-12-30 First_np_Paper/first_np_paper.tex
Updated section 2.3 with first round edits

\author[geo,eps]{D.XXX. Rutte}

@@

\comment{Need Daniel's middle initial}

@@

The kinematics of D-D fusion produce neutrons of energy ranging from 2.2 to 2.85 MeV, over the lab frame emission angular range of 0-180\degree, with respect to the incident deuteron beam. This distribution is a function of the incident deuteron kinetic energy, and has been well-documented; the energy-angle distribution for the incident 100 keV deuterons used in the HFNG can be seen plotted in \autoref{fig:scatt_angle}.

@@

Collimated beamlines can be used to select an energy group from this distribution, by permitting only neutrons within a certain angular window to be incident upon a target. Due to the compact nature of the HFNG, this would require the mounting of samples outside of the target chamber, with at least several feet of standoff from the D-D reaction surface. This would greatly reduce the flux (to approximately 10$^4$ neutrons/cm$^2$ over 4$\pi$) subtended by a single target and require longer irradiations for activation, which exacerbates the effect of simultaneous production \& decay in activation measurements. Instead, samples are mounted internally in the HFNG target chamber, separated by 5 mm of copper from the D-D reaction surface. This allows the samples to subtend a fairly narrow (28\degree) angle of the forward-focused neutrons, while maintaining the high (approximately 10$^8$ neutrons/cm$^2$ ) flux for activations.

@@

To determine the energy distribution of neutrons incident upon the loaded foils, MCNP6 was used to model the generator operation, and produced a neutron energy spectrum incident upon target foils co-loaded into the HFNG, seen in \autoref{fig:mcnp_flux}. This spectrum verifies the forward-focused kinematics of the D-D reaction subtended by the co-loaded sample foils.

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/mcnp_flux.pdf}
 % mcnp_flux.pdf: 373x295 pixel, 72dpi, 13.16x10.41 cm, bb=0 0 373 295
 \caption{MCNP6-modeled neutron energy spectrum for the Berkeley HFNG, using co-loaded indium and zinc (for \ce{^{64}Cu} production) foils as an example.}
 \label{fig:mcnp_flux}
\end{figure}

While this shows that the sample foils receive a very narrow energy distribution of incident neutrons, an effective neutron energy and energy bounds must be determined to constrain cross section ratio measurements. From the same MCNP6 simulation, flux tallies show a flux-weighted average neutron energy of 2.7645 MeV in the Indium foil, and average neutron energy of 2.7649 MeV in the zinc or titanium foil; these are indistinguishable within statistical uncertainty, and serve as an effective neutron energy in the loaded target foils. Due to the kinematics of D-D neutron emission, $E_{max}$,  the maximum energy of a neutron subtending the target foils in this geometry, can be calculated as 2.7825 MeV. For this maximum energy, the number of reactions induced in a foil (containing $N_i$ target nuclei) is simply:

@@

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/fracplot.pdf}
 % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
 \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]}
 \label{fig:frac_plot}
\end{figure}

This quantity $F\pp{E'}$ is plotted in \autoref{fig:frac_plot}. Using the flux-averaged neutron energy in the Indium foil as a centroid, this fraction of total reactions can be used to to estimate the uncertainty in effective neutron energy, which will become energy error bars in the excitation function for these reactions. Within 1 standard deviation above and below the centroid, 68.2\% of all reactions should be encompassed. Thus, the upper and lower energy uncertainties correspond to the neutron energies $E'$  such that $F\pp{E'}$ = 0.841 and $F\pp{E'}$ = 0.159, respectively. From \autoref{eqn:react_fraction},  these are found to be  F(2.7796 MeV) = 0.841 and F(2.7426 MeV) = 0.159, respectively; thus the effective neutron energy subtended by target foils is reported as $E_n$ = 2.7645 MeV + 0.0151 MeV / - 0.0219 MeV.  This 37-keV full-energy spread verifies that, at such close distances to the D-D reaction surface, loaded target foils receive a highly quasi-monoenergetic neutron flux.

@@

@@

>>>>

\author[geo,eps]{D. Rutte}

@@

% \comment{Need Daniel's middle initial}

@@

The kinematics of D-D fusion at 100 keV lab energy produce neutrons of energy ranging from 2.2 to 2.85 MeV, over the lab-frame emission angular range of 0-180\degree, with respect to the incident deuteron beam. This distribution is a function of the incident deuteron kinetic energy, and has been well-documented \cite{Liskien_Paulsen_1973}; \autoref{fig:scatt_angle} displays the energy-angle distribution for the incident 100 keV deuterons used in the HFNG.

@@

Collimated beamlines can be used to select an energy group from this distribution, by permitting only neutrons within a certain angular window to be incident upon a target. Due to the compact nature of the HFNG, this would require the mounting of samples outside of the target chamber, with a distance of at least 20 cm from the D-D reaction surface. This would greatly reduce the flux (to approximately 10$^4$ neutrons/s\ cm$^2$ over 4$\pi$) subtended by a single target and require longer irradiations for activation, which exacerbates the effect of simultaneous production and decay in activation measurements. Instead, the samples are mounted internally in the HFNG target chamber, separated by 5 mm of copper from the D-D reaction surface. This allows the samples to subtend a fairly narrow ($\sim$28\degree) angle of the forward-focused neutrons, while maintaining the high (approximately $\sci{1.3}{7}$neutrons/s\ cm$^2$ ) flux for activations.

\comment{Make global flux / fluence unification}

@@

\comment{Need to replace this figure - coming from Mauricio?}

To determine the energy distribution of neutrons incident upon the loaded foils following production and transport, the Monte Carlo N-Particle transport code  MCNP6 \cite{goorley2013initial} was used to model the neutron energy spectrum incident upon target foils co-loaded into the HFNG (see \autoref{fig:mcnp_flux}). This spectrum illustrates the forward-focused kinematics of the D-D reaction subtended by the co-loaded sample foils.

% \begin{figure}
%  \centering
%  \includegraphics[scale=0.6]{./figures/mcnp_flux.pdf}
%  % mcnp_flux.pdf: 373x295 pixel, 72dpi, 13.16x10.41 cm, bb=0 0 373 295
%  \caption{MCNP6-modeled neutron energy spectrum for the Berkeley HFNG, using co-loaded indium and zinc (for \ce{^{64}Cu} production) foils as an example.}
%  \label{fig:mcnp_flux}
% \end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/mcnp_flux_new.pdf}
 % mcnp_flux.pdf: 373x295 pixel, 72dpi, 13.16x10.41 cm, bb=0 0 373 295
 \caption{MCNP6-modeled neutron energy spectrum for the Berkeley HFNG, using co-loaded indium and zinc (for \ce{^{64}Cu} production) foils as an example.}
 \label{fig:mcnp_flux}
\end{figure}

\comment{Use this version, old version (mcnp-flux.pdf), or some new combination of lines?}

While this shows that the sample foils receive a very narrow energy distribution of incident neutrons, an effective neutron energy and energy bounds must be determined to constrain cross section ratio measurements. From the same MCNP6 simulation, flux tallies show a flux-weighted average neutron energy of 2.7645 MeV in the Indium foil, and 2.7649 MeV in the zinc or titanium foil; these are indistinguishable within statistical uncertainty, and serve as an effective neutron energy in the loaded target foils. Due to the kinematics of D-D neutron emission, $E_{max}$,  the maximum energy of a neutron subtending the target foils in this geometry is 2.7825 MeV. For this maximum energy, the number of reactions induced in a foil (containing $N_i$ target nuclei) is simply:

@@

% \begin{figure}
%  \centering
%  \includegraphics[scale=0.6]{./figures/fracplot.pdf}
%  % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
%  \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]}
%  \label{fig:frac_plot}
% \end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/fracplot_new.pdf}
 % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
 \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]}
 \label{fig:frac_plot}
\end{figure}

\comment{Use this version, or old version (fracplot.pdf)?}

This quantity $F\pp{E'}$ is plotted in \autoref{fig:frac_plot}. Using the flux-averaged neutron energy in the Indium foil as a centroid, this fraction of total reactions can be used to estimate the uncertainty in effective neutron energy, which will become energy error bars in the excitation function for these reactions. Within 1 standard deviation above and below the centroid, 68.2\% of all reactions should be encompassed. Thus, the upper and lower energy uncertainties correspond to the neutron energies $E'$  such that $F\pp{E'}$ = 0.841 and $F\pp{E'}$ = 0.159, respectively. From \autoref{eqn:react_fraction},  these are found to be  F(2.7796 MeV) = 0.841 and F(2.7426 MeV) = 0.159, respectively; thus the effective neutron energy subtended by target foils is reported as $E_n$ = 2.7645 MeV + 0.0151 MeV / - 0.0219 MeV.  This 37-keV full-energy spread verifies that, at such close distances to the D-D reaction surface, loaded target foils receive a highly quasi-monoenergetic neutron flux.

@@

\comment{Does \autoref{fig:main_ge_detectors} serve any purpose? Should it be cut?}

@@

\comment{Need to update \autoref{fig:decay_curves} - figures are blurry}
